Relative Stable Maps in Genus One via Central Alignments

### 2018-12-03 new.tex
started combinatorial and explicit description

\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\VZ}{\pazocal{V\!Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\textbf{Gathmann-type recursion.} There is a forgetful morphism \[\VZ_{1,\alpha}(\PP^N|H,d)\to \VZ_{1,n}(\PP^N,d),\] hitting Gathmann's relative space. We may therefore pullback Gathmann's line bundle and section, cutting out the locus where the $k$-th marking is tangent to $H$ to order $\alpha_k+1$. Because $\alpha$ was maximal ($\sum\alpha=d$) by assumption, this means that the curve has to break, and $x_k$ has to lie on an internal component - one which is entirely mapped into $H$. We identify the zero locus of Gathmann's section explicitly. Here is an interesting remark: the combinatorics of such boundary loci is governed by tropical geometry, and it is otherwise very hard to districate the interaction between the relative condition and the exceptional loci of the Vakil--Zinger blow-up.
\begin{cor}
 $\VZ_{1,\alpha}(\PP^N|H,d)$ is smooth over its Artin fan, in particular codimension one logarithmic strata can be read off from the latter.\marginpar{why are we only interested in log strata?}
\end{cor}
The Artin fan is a tropical gadget. Its local structure is given by subdividing the ACGS minimal monoid according to the alignment. We are only interested in picking its rays. The upshot is that the combinatorics is slightly more involved than in the genus zero case: the alignment may force some teeth of the comb to break.
\begin{thm}[Gathmann-type formula, maximal tangency, $(\PP^N|H)$ case]
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{1,\alpha;k}(\PP^N|H,d)],\]
 the latter being a sum of broken comb loci indexed by rays of the tropical fan.
\end{thm}
Importantly, the broken comb loci admit a very explicit description in terms of tautological integrals on the underlying boundary of Gathmann's relative space.
\begin{thm}
 Up to a finite cover of the underlying boundary stratum - which is a combinatorially-determined fiber product of moduli spaces of genus zero and one, absolute and relative maps with lower numerical invariants - every component of $D_{\alpha,k}(\PP^N|H,d)$ can be described as the transverse intersection of two loci in a projective bundle, where:
 \begin{itemize}
  \item the latter parametrises the possible line bundle isomorphisms imposed by the alignment of the log structure;
  \item the first locus is a subbundle representing the residual isomorphisms after fixing the ones virtually imposed by tropical continuity;
  \item the second locus is determined by the factorisation conditions.
 \end{itemize}
\end{thm}
The upshot is that we may then push the formula down to the Gathmann's space, so as to obtain multiplicities and tautological classes.
\begin{cor}
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ^G_{1,\alpha}(\PP^N|H,d)]=[D^G_{1,\alpha;k}(\PP^N|H,d)],\]
 the latter being expressible as a weighted sum of tautological classes on Gathmann's comb loci.
\end{cor}
Once we have this formula, the following extensions are classical:
\begin{itemize}
 \item A similar formula for raising the tangency holds in the non-maximal tangency case. It can be proven by adding auxiliary markings of contact order $1$; forgetting them is then a $(d-\sum\alpha)!:1$ cover because the nice locus is dense inside Gathmann's relative spaces.
 \item The formula holds more generally for any smooth projective target $X$ relative to a generic hyperplane section $Y=X\cap H\subseteq \PP^N$. This follows via virtual pullback.
\end{itemize}
Finally the recursive structure of the boundary allows us to prove the following
\begin{thm}[In-principle quantum Lefschetz]
 The restricted reduced genus one invariants of $Y$ can be inductively deducted from the full descendant genus zero and one (reduced) Gromov-Witten theory of $X$.
\end{thm}
The proof is more delicate than its genus zero analogue because invariants with the same numerical data appear intertwined in the last steps of the recursion.

@@

Note that $f$ has contact order with $H$ exactly equal to $\alpha_k$ at the marking $p_k$, or else the irreducible component of $C$ on which $x_k$ lies is mapped entirely inside $H$. By pulling back along $f$ the equation defining $H$, and taking its $\alpha_k+1$-st derivative at $x_k$ (which makes sense because all the lower order derivatives do vanish by assumption) we single out the latter locus. By staring at the exact sequence of jet bundles
\[0\to p_k^*\Omega_{C}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to p_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to p_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0\]
we realise that there is on $\VZ_{1,\alpha}(\PP^N|H,d)$ a line bundle (the leftmost term in the exact sequence) with a natural section cutting the locus where the curve breaks, and the piece containing $p_k$ is mapped inside $H$.
\begin{lem}
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{1,\alpha;k}(\PP^N|H,d)],\]
\end{lem}
In what follows we shall give an explicit (recursive) description of the right-hand side, in terms of tautological classes on spaces of maps with lower numerical invariants.
\begin{remark}
 The divisors we are after are a union of logarithmic strata, because the locus where the log structure is trivial coincides precisely with the nice locus.
\end{remark}
By Corollary \ref{cor:log_smooth}, the divisorial logarithmic strata correspond to rays of the Artin fan.

\subsection{Combinatorial description}
As long as the minimal subcurve of genus one has positive degree, the picture is classical: the dual graph of the curve is bipartite - vertices mapped inside $H$ are distinguished from the others. In fact there is only one such vertex, because, if there were more, the relative position of their image in $\mathbb R_{\geq 0}$ under the tropical map would be unconstrained, therefore spanning a locus larger than a ray in the tropical moduli space. We follow in R. Vakil's footsteps in setting the following notation:
\begin{figure}[h]
    \centering
    \begin{minipage}{0.3\textwidth}
        \centering
        \Yagraph
        \caption{$\mathcal Y_a$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ybgraph
        \caption{$\mathcal Y_b$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ycgraph
        \caption{$\mathcal Y_c^+$}
    \end{minipage}
\end{figure}

@@

On the other hand, when there is a contracted elliptic subcurve - and it will be contracted into the hyperplane, because otherwise it wouldn't be generic, by density of the nice locus in $\VZ_{1,\tilde\alpha}(\PP^N|H,\tilde d)$ - the picture becomes much more complicated due to the alignment. The combs may break multiple times and they may even point in opposite directions.
\begin{figure}
 \tikz{
 \draw (0,0) circle(2pt)[fill=black] (0,2) circle(2pt)[fill=black] (0,4) circle(2pt)[fill=black] (0,6) circle(2pt)[fill=black] (0,-2) circle(2pt)[fill=black] (2,0) circle(2pt)[fill=white] (1.1,1.78) circle(2pt)[fill=black] (3.42,1.42) circle(2pt)[fill=black] (.58,-1.42) circle(2pt)[fill=black];
 \draw (2,0) -- node[above]{4} (0,0) (2,0) --node[above]{2} (1.1,1.78) --node[above]{2} (0,2) (1.1,1.78) --node[above]{1} (0,4) (2,0) --node[above]{1} (3.42,1.42) (2,0) --node[above]{1} (.58,-1.42) (3.42,1.42) --node[above]{5} (0,6) (.58,-1.42) --node[above]{3} (0,-2);
 \draw[red] (2,0) circle(2);
 }
\end{figure}

\subsection{Boundary loci}
We follow in Vakil's footsteps in setting the following notation:
\begin{enumerate}[label=(\alph*)]
 \item \[\mathcal Y^a=\M{0}{\lvert\alpha^{(0)}\rvert+r}{H}{d_0}\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y^b=\M{0}{\lvert\alpha^{(0)}\rvert+r}{H}{d_0}\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y^c\subseteq \VZc{1}{\lvert\alpha^{(0)}\rvert+r}{H}{d_0}\times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
 $\mathcal Y^c$ is cut within the latter by Gathmann's condition on line bundles:
 \[f_{|E}^*\mathcal O_H(1)\cong\mathcal O_E\left(\sum_{x_j\in E}\alpha_jx_j-\sum_{i=1}^r m^{(i)}y_i\right)\]\sidenote{Express the line bundle condition in terms of tautological classes?}
 Notice that such condition depends on the choice of $m^{(i)}$ but is really just a condition on the first factor of the product, so $\mathcal Y^c$ itself can be expressed as a product.
\end{enumerate}

>>>>

\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\MG}[4]{\overline{\mathcal{M}}^{\rm G}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\VZ}{\pazocal{V\!Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\fgt}{\operatorname{fgt}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\textbf{Gathmann-type recursion.} There is a forgetful morphism \[\VZ_{1,\alpha}(\PP^N|H,d)\to \VZ_{1,n}(\PP^N,d),\] hitting Gathmann's relative space. We may therefore pullback Gathmann's line bundle and section, cutting out the locus where the $k$-th marking is tangent to $H$ to order $\alpha_k+1$. Because $\alpha$ was maximal ($\sum\alpha=d$) by assumption, this means that the curve has to break, and $x_k$ has to lie on an internal component - one which is entirely mapped into $H$. We identify the zero locus of Gathmann's section explicitly. Here is an interesting remark: the combinatorics of such boundary loci is governed by tropical geometry, and it is otherwise very hard to districate the interaction between the relative condition and the exceptional loci of the Vakil--Zinger blow-up.
\begin{cor}
 $\VZ_{1,\alpha}(\PP^N|H,d)$ is smooth over its Artin fan, in particular codimension one logarithmic strata can be read off from the latter.
\end{cor}
The Artin fan is a tropical gadget. Its local structure is given by subdividing the ACGS minimal monoid according to the alignment. We are only interested in picking its rays. The upshot is that the combinatorics is slightly more involved than in the genus zero case: the alignment may force some teeth of the comb to break.
\begin{thm}[Gathmann-type formula, maximal tangency, $(\PP^N|H)$ case]
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{1,\alpha;k}(\PP^N|H,d)],\]
 the latter being a sum of broken comb loci indexed by rays of the tropical fan.
\end{thm}
Importantly, the broken comb loci admit a very explicit description in terms of tautological integrals on the underlying boundary of Gathmann's relative space.
\begin{thm}
 Up to a finite cover of the underlying boundary stratum - which is a combinatorially-determined fiber product of moduli spaces of genus zero and one, absolute and relative maps with lower numerical invariants - every component of $D_{\alpha,k}(\PP^N|H,d)$ can be described as the transverse intersection of two loci in a projective bundle, where:
 \begin{itemize}
  \item the latter parametrises the possible line bundle isomorphisms imposed by the alignment of the log structure;
  \item the first locus is a subbundle representing the residual isomorphisms after fixing the ones virtually imposed by tropical continuity;
  \item the second locus is determined by the factorisation conditions.
 \end{itemize}
\end{thm}
The upshot is that we may then push the formula down to the Gathmann's space, so as to obtain multiplicities and tautological classes.
\begin{cor}
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ^G_{1,\alpha}(\PP^N|H,d)]=[D^G_{1,\alpha;k}(\PP^N|H,d)],\]
 the latter being expressible as a weighted sum of tautological classes on Gathmann's comb loci.
\end{cor}
Once we have this formula, the following extensions are classical:
\begin{itemize}
 \item A similar formula for raising the tangency holds in the non-maximal tangency case. It can be proven by adding auxiliary markings of contact order $1$; forgetting them is then a $(d-\sum\alpha)!:1$ cover because the nice locus is dense inside Gathmann's relative spaces.
 \item The formula holds more generally for any smooth projective target $X$ relative to a generic hyperplane section $Y=X\cap H\subseteq \PP^N$. This follows via virtual pullback.
\end{itemize}
Finally the recursive structure of the boundary allows us to prove the following
\begin{thm}[In-principle quantum Lefschetz]
 The restricted reduced genus one invariants of $Y$ can be inductively deducted from the full descendant genus zero and one (reduced) Gromov-Witten theory of $X$.
\end{thm}
The proof is more delicate than its genus zero analogue because invariants with the same numerical data appear intertwined in the last steps of the recursion.

@@

Note that $f$ has contact order with $H$ exactly equal to $\alpha_k$ at the marking $p_k$, or else the irreducible component of $C$ on which $p_k$ lies is mapped entirely inside $H$. By pulling back along $f$ the equation defining $H$, and taking its $\alpha_k$-th derivative at $p_k$ (which makes sense because all the lower order derivatives do vanish by assumption) we single out the latter locus. By staring at the exact sequence of jet bundles
\[0\to p_k^*\Omega_{C}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to p_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to p_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0\]
we realise that there is on $\VZ_{1,\alpha}(\PP^N|H,d)$ a line bundle (the leftmost term in the exact sequence above) with a natural section cutting the locus where the curve breaks, and the piece containing $p_k$ is mapped inside $H$.
\begin{lem}
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{1,\alpha;k}(\PP^N|H,d)],\]
\end{lem}
In what follows we shall give an explicit (recursive) description of the right-hand side, in terms of tautological classes on spaces of maps with lower numerical invariants.
\begin{remark}
 The divisors we are after are a union of logarithmic strata, because the locus where the log structure is trivial coincides precisely with the nice locus.
\end{remark}
By Corollary \ref{cor:log_smooth}, the logarithmic strata of codimension one correspond to rays of the Artin fan.

\subsection{Combinatorial description}
As long as the minimal subcurve of genus one has positive degree, the picture is classical: the dual graph of the curve is bipartite - vertices mapped inside $H$ are distinguished from the others. In fact there is only one such vertex, because, if there were more, the relative position of their image in $\mathbb R_{\geq 0}$ under the tropical map would be unconstrained, therefore spanning a locus larger than a ray in the tropical moduli space. \marginpar{say something about rigid tropical curves as in ACGS and KLR} We follow in R. Vakil's footsteps in setting the following notation:
\begin{figure}[h]
    \centering
    \begin{minipage}{0.3\textwidth}
        \centering
        \Yagraph
        \caption{$\mathcal Y_a$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ybgraph
        \caption{$\mathcal Y_b$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ycgraph
        \caption{$\mathcal Y_c^+$}
    \end{minipage}
\end{figure}

@@

On the other hand, when there is a contracted elliptic subcurve - and it will be contracted into the hyperplane, because otherwise it wouldn't be generic, by density of the nice locus in $\VZ_{1,\alpha^\prime}(\PP^N|H,d^\prime)$ - the picture becomes more complicated due to the alignment. The combs may break (multiple times) and they may point in opposite directions. We label these loci $\mathcal{Y}_c^0$.
\begin{figure}
 \tikz{
 \draw (0,0) circle(2pt)[fill=black] (0,2) circle(2pt)[fill=black] (0,4) circle(2pt)[fill=black] (0,6) circle(2pt)[fill=black] (0,-2) circle(2pt)[fill=black] (1.1,1.78) circle(2pt)[fill=black] (3.42,1.42) circle(2pt)[fill=black] (.58,-1.42) circle(2pt)[fill=black];
 \draw (2,0) -- node[above]{4} (0,0) (2,0) --node[above]{2} (1.1,1.78) --node[above]{2} (0,2) (1.1,1.78) --node[above]{1} (0,4) (2,0) --node[above]{1} (3.42,1.42) (2,0) --node[above]{1} (.58,-1.42) (3.42,1.42) --node[above]{5} (0,6) (.58,-1.42) --node[above]{3} (0,-2);
 \draw[red] (2,0) circle(2);
 \draw (2,0) circle(2pt)[fill=white];
 \draw[->] (0,-3) circle(2pt)[fill=black] edge (4,-3);
 }
\end{figure}
\begin{dfn}
 A degree-aligned genus one tropical map $\phi$ to $\mathbb R_{\geq 0}$ is ... with a circle of radius $\delta$ around the core satisfying:
 \begin{enumerate}
  \item the circle of radius $\delta$ passes through at least one vertex of positive degree, and no vertex of positive degree is contained in its strict interior;
  \item every circle of radius $<\delta$ around the core passes through (either none or) at least two vertices;\footnote{The subdivisions we perform before picking a ray correspond to aligning the log structure, i.e. imposing $\lambda(v)=\lambda(v^\prime)$ for some $v,v^\prime$.}
  \item every edge heading out of the circle of radius $\delta$ goes directly to a vertex in $\phi^{-1}(0)$.
 \end{enumerate}
\end{dfn}

\begin{remark}
 The core being contracted in the fiber of the tropical map is not a phenomenon that we should worry about in codimension one. Indeed, assume that the core is contracted in the fiber along a ray. Then all the edges departing from the core have expansion factor $0$; call the corresponding coordinates $U=\{u_i\}_{i\in I}$. Call the remaining coordinates $E=\{e_j\}_{j\in J}$. Note that tropical continuity involves only $E$. Alignments on the other hand assume the form $\lambda(v)=\lambda(v^\prime)$, where $\lambda(v)=\sum_{i\in I(v)}u_i\sum_{j\in J(v)}e_j$. Pick the shortest elements of $U$; then these can be shortened to zero without affecting the rest (by hypothesis, alignments can only identify them among themselves). This shows that we could not have started with a ray.
\end{remark}

\subsection{Explicit realisation}
The description of the first three types is relatively easy:
\begin{enumerate}[label=(\alph*)]
 \item \[\mathcal Y_a=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_b=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_c^+= \VZc{1}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^{\sim} \times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
\end{enumerate}
where $\M{0}{\mu}{H}{d}^\sim$ denotes the space of genus zero punctured maps, while $\VZc{0}{\mu}{H}{d}$ denotes the desingularisation of the main component of the space of punctured maps discussed in Section ???. The numerical data run over all splitting $A$ of the tangency conditions (resp. $B$ of the degree) in $r+1$ parts, such that:
\begin{enumerate}
 \item $d_0+\sum_{i=1}^r m^{(i)}=\sum\alpha^{(0)}$;
 \item $d_i=m^{(i)}+\sum\alpha^{(i)}$;
 \item $d_0>0$ in case $c$.
\end{enumerate}

Review here the process of gluing for punctured maps. In particular the basic monoid should be modified in order to make the relevant evaluations log morphisms. Claim: evaluations are strict. Consequence: the fiber product in the category of log stacks is fine. We have then to apply saturation. This is a finite morphism of degree... (I think it could be $\frac{\prod m^{(i)}}{lcm(m^{(i)}}$).

\begin{lemma}[Virtual pushforward] The following hold.
 \begin{itemize}
  \item $\fgt_*[\M{0}{\alpha}{\PP^N|H}{d}]=[\MG{0}{\alpha}{\PP^N|H}{d}]$ (follows from \cite{GathmannThesis,AbramovichMarcusWiseComparison}).
  \item $\fgt_*[\VZc{1}{\alpha}{\PP^N|H}{d}]$ computes the reduced relative invariants by definition.
  \item $\fgt_*[\M{0}{\mu}{H}{d_0}^\sim]=[\M{0}{\lvert\mu\rvert}{H}{d_0}]$ (follows from \cite{GathmannThesis} and... comparison of punctured with rubber invariants).
  \item $\fgt_*[\VZc{1}{\mu}{H}{d_0}^\sim]$ here we should need a variation on Pixton's DRC formula; hopefully it's enough to avoid the graphs that tropical well-spacedness discards.
 \end{itemize}
\end{lemma}
